---
title: "Diary Entries NM2207"
author: "Ariel Quek"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Diary Entry #1

**(1) Finalized Topic for Final Project:**

Premier League matches from 2021-2022 season!

In the Premier League 2021-2022, there were many surprises undoubtedly, and the predictions made during the season varied. There have been many debate about outstanding players as well, with many critics/commentators seeming to agree only on Kevin De Bruyne from Manchester City.

It seems interesting to look into the matches leading up to Man City's win, while investigating the consistency of the various club's play on the field.

**(2) Data sources curated so far:**

*Data sources are derived from [[Tidy tuesday](%5Bhttps://github.com/rfordatascience/tidytuesday#about-tidytuesday)](https://github.com/rfordatascience/tidytuesday#about-tidytuesday), as recommended via Canvas!*

The CSV files are downloaded in my local desktop:

\- soccer 21-22.csv

\- weeklyrank.csv

The links to relevant data sets and websites can be found here:

\- <https://github.com/rfordatascience/tidytuesday#about-tidytuesday>

\- <https://www.kaggle.com/datasets/evangower/premier-league-match-data>

\- end -

## Diary Entry #2

### Question 1:

***Finalised Idea/ Theme*****:**

-   Topic: A quantitative analysis of the 20/21 Premier League season

-   Question: "**How influential was the home advantage for the football clubs in the 20-21 Premier League Season?"**

-   *Considering the performance analysis and the relationship with the match outcomes*

**In order to answer the above question:**

-   First, look at the performance analysis by calculating and comparing the average number of points and goal differences for each team over the season.

-   Second, look at the percentage of matches won by the home team and the away team, and compare.

-   Lastly, look at the correlation between the two, especially looking at the correlation between perhaps the number of shots on target for home teams as opposed to away teams, etc.

### Question 2

-   This is an important question because it is commonly assumed that football teams are more likely to win if they're playing at home. However, the best way to answer this question accurately is through data analysis, as it allows us to use the precise data of goals scores by each time to measure the home advantage.

-   This can be helpful in predicting the scores for future football matches, and identifying how relevant the fan atmosphere is for matches. This information can be very influential considering the post-Covid boom of spectators for various sports.

-   On more personal fronts, as a football fan, this question will help with match analysis, especially with the ongoing Premier League season.

## Data:

### Question 3

-   The data packages are as follows:

Downloading data package:

```{r}
#Opening the relevant CSV files   

##Reading soccer data from 21-22 Premier League

soccer <- read.csv("soccer21-22.csv")
head(soccer)
```

```{r}
## Reading the weekly rankings of the FCs  

weeklyrank <- read.csv("weeklyrank.csv") 
head(weeklyrank)
```

From these data packages, the following columns and rows will be used:

-   In "weeklyrank", I will be using the columns "Team", "Rank" and "Week"

-   In "soccer", I will be using "HomeTeam", "AwayTeam", "FTHG", "FTAG", "FTR", "HTHG", "HTAG", "HTR", "HS", "AS", "HST", "AST".

-   <div>

    Clarifying acronyms

    -   FTHG - full time home goals

    -   FTAG - full time away goals

    -   HTHG - half time home goals

    -   HTAG - half time away goals

    -   FTR - full time result

    -   HTR - half time result

    -   HS - home shots

    -   AS - away shots

    -   HST - home shots on target

    -   AST - away shots on target

    </div>

\- end -

## Diary Entry #3

***Visualizations for final project:***

-   Bar charts: Plotting the wins of each team when on home grounds, and plotting a separate graph for wins on away grounds.

<!-- -->

-   Scatter plots: Plotting weekly rankings for teams on home grounds as opposed to on away grounds.

-   Pie charts: To plot home advantage influence for each team

***How I plan to make it interactive:***

-   ggplot2 package: To plot bar charts and scatter plots colour coded for each team

shiny package: For addition of app-like elements (whether through sliders or inputs to allow users to see specific data for specific games or filter for specific teams they are fans of)

***Concepts incorporated in my final project:***

+------------+------------------------------------------------------------------------------+
| Week(s)    | Topics/Concepts                                                              |
+============+==============================================================================+
| 1          | R Markdown concepts:                                                         |
|            |                                                                              |
|            | -   Adding photos, changing the appearance of text (**bold**, *italicizing*) |
+------------+------------------------------------------------------------------------------+
| 2          | Data Visualization concepts:                                                 |
|            |                                                                              |
|            | -   ggplot2 package                                                          |
|            |                                                                              |
|            | -   Plotting scatter plots                                                   |
|            |                                                                              |
|            | -   Plotting bar graphs                                                      |
+------------+------------------------------------------------------------------------------+
| (self)     | Data Visualization: Pie chart                                                |
+------------+------------------------------------------------------------------------------+
| 7          | Data Visualization:                                                          |
|            |                                                                              |
|            | -   Visualizing different types of variables using ggplot2                   |
+------------+------------------------------------------------------------------------------+
| 8          | Data Visualization:                                                          |
|            |                                                                              |
|            | -   Creating interactive elements using Shiny and Quarto                     |
+------------+------------------------------------------------------------------------------+

\- end -

## Diary Entry #4 

After starting on the data analysis and visualization, some challenges encountered involved the scope and perimeters I had set for the project at the start:

**Instead of doing data analysis for all 20 teams in the EPL, I narrowed it down to 6**

-   This is because the previous scope for analysis was too broad and the data could not be represented well visually

-   With just 6 teams instead, the data can be represented more evenly and better.

-   Using the 6 teams as a sample is also viable considering the history of their performance in the EPL.

-   This would be better than relying on the big 4, as it gives more teams' data to look at.

\- end -
