---
title: "Challenge 7"
author: "Ariel Quek"
date: "`r Sys.Date()`"
output: 
  pdf_document: default
  html_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Challenge 7 Instructions 

1. You are required to focus on the lecture videos and materials, owing to the break last week

2. The material taught this week is of significant importance for your final projects

3. The challenge this week will be your workbook: a document that you will generate by trying and experimenting with the code snippets taught in the lecture material.

```{r, echo=TRUE,eval=TRUE}
library(tidyverse)
install.packages("palmerpenguins",repos = "http://cran.us.r-project.org")
library(palmerpenguins)
View(penguins)
glimpse(penguins)
```

## Recreating Plot for Palmer Penguins 

Plotting the bill depth against bill length of penguins 

```{r, echo=TRUE, eval=TRUE}
ggplot(data = penguins,
       mapping = aes(x = bill_depth_mm,
                     y = bill_length_mm,
                     colour = species))+
  geom_point() +
  labs(title = "Bill depth and length", 
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x ="Bill depth (mm)", 
       y = "Bill length (mm)",
       colour = "Species",
       caption = "Source: Palmer Station LTER/ palmerpenguins package",
scale_colour_viridis_d())
```
Plotting bill depth against bill length with shaped data points: 

```{r, echo=TRUE, eval=TRUE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, colour = species,
shape = island)) +
geom_point() + scale_colour_viridis_d()
```

Changing the size of data points on the plot 

```{r, eval=TRUE, echo=TRUE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, colour = species, shape = species,
size = body_mass_g)) +
geom_point() + scale_colour_viridis_d()
```

Manipulating alpha (opacity) of data points in the plot 

```{r, eval=TRUE, echo=TRUE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, colour = species,
shape = species, size = body_mass_g, alpha = flipper_length_mm)) +
geom_point() + scale_colour_viridis_d()
```

## Faceting: creating smaller plots that display different subsets of the data 

*1. Plotting the bill length and bill depth of penguins based on the species and the island* 
```{r, echo=TRUE, eval=TRUE}
ggplot(penguins) +
aes(x = bill_depth_mm,
y = bill_length_mm) +
geom_point() +
facet_grid(species ~ island)
```

*2. Plotting bill length and bill depth of penguins based on sex against species* 
```{r, eval=TRUE, echo=TRUE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point() +
facet_grid(species ~ sex)
```

*3. Alternatively, plotting bill length and bill depth of penguins based on species against sex* 
```{r, echo=TRUE, eval=TRUE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point() +
facet_grid(sex ~ species)
```

*4. Plotting bill length against bill depth only by species* 
```{r, echo=TRUE, eval=TRUE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point() +
facet_wrap(~ species)
```
*5. Plotting bill length against bill depth by species, in separate grids*
```{r, echo=TRUE, eval=TRUE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + geom_point()+
facet_wrap(~ species, ncol =2)
```
*6. Adding colour (without a legend)* 
```{r, echo=TRUE, eval=TRUE}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm, color = species)) +
geom_point() + facet_grid(species ~ sex) + scale_color_viridis_d() +
guides(color =
"none"
)
```
## Visualizing numeric variables using Lending Club Data set

Retrieving dataset
```{r, echo=TRUE, eval=TRUE}
install.packages("openintro", repos = "http://cran.us.r-project.org")
library(openintro)
library(tidyverse)
glimpse(loans_full_schema)
```
Selecting variables
```{r, echo=TRUE, eval=TRUE}
loans <- loans_full_schema %>%
select(loan_amount, interest_rate, term, grade,
state, annual_income, homeownership, debt_to_income)
glimpse(loans)
```
Plotting histogram of loan amount against loan count with different binwidth lengths 

*1. binwidth = 1000*
```{r, echo=TRUE, eval=TRUE}
ggplot(loans) + aes(x = loan_amount) +
geom_histogram(binwidth = 1000)
```

*2. binwidth = 5000*
```{r, echo=TRUE, eval=TRUE}
ggplot(loans) + aes(x = loan_amount) +
geom_histogram(binwidth = 5000)
```

*3. binwidth = 20000* 
```{r, echo=TRUE, eval=TRUE}
ggplot(loans) + aes(x = loan_amount) +
geom_histogram(binwidth = 20000)
```

Customizing histograms, with colors and labels
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) +
geom_histogram(binwidth = 5000, alpha = 0.5) + labs(x ="Loan amount ($)",y ="Frequency",title ="Amounts of Lending Club loans")
```

Creating facet plot of histograms with categorical variable 
```{r, echo=TRUE, eval=TRUE}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) + geom_histogram(binwidth = 5000) + labs(x ="Loan amount ($)",y =
"Frequency",title ="Amounts of Lending Club loans") +facet_wrap(~ homeownership, nrow =3)
```

Creating Density Plot with customizable bandwidth

1. Default bandwidth 
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = loan_amount)) +
geom_density(adjust=1)
```

2. Increased bandwidth (2)
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = loan_amount)) +
geom_density(adjust=2)
```

3. Decreased bandwidth (0.5)
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = loan_amount)) +
geom_density(adjust=0.5)
```

Customizing density plot!
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = loan_amount, fill = homeownership)) +
  geom_density(adjust =2, alpha =0.5) +labs(x ="Loan amount ($)",y ="Density",title ="Amounts of Lending Club loans", fill ="Homeownership")
```
## Plotting box plots 

Plotting interest rate of loans 
```{r, echo=TRUE, eval=TRUE}
ggplot(loans, aes(x = interest_rate)) +geom_boxplot()
```

Plotting annual income against loans 
```{r, echo=TRUE, eval=TRUE}
ggplot(loans, aes(x = annual_income)) +
geom_boxplot()
```
Customizing box plots 
```{r, echo=TRUE, eval=TRUE}
ggplot(loans, aes(x = interest_rate)) +geom_boxplot() +labs(x ="Interest rate (%)",y =NULL,title ="Interest rates of Lending Club loans") +theme( axis.ticks.y = element_blank(), axis.text.y = element_blank() )
```

By grade of loan 
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = interest_rate,y = grade)) +geom_boxplot() +labs(x =
"Interest rate (%)",y ="Grade",title ="Interest rates of Lending Club loans",subtitle= "by grade of loan")
```

## Creating Scatterplot 
```{r, echo=TRUE, eval=TRUE}
ggplot(loans, aes(x = debt_to_income, y = interest_rate)) +
geom_point()
```

## Creating Hex Plot 
```{r, echo=TRUE, eval=TRUE}
ggplot(loans %>% filter(debt_to_income <100),aes(x = debt_to_income, y = interest_rate)) +geom_hex()
```

## Creating Bar Plot 
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = homeownership)) +geom_bar()
```

Customizing segmented bar plot (by grade of homeownership)
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = homeownership,
fill = grade)) +
geom_bar()
```

Customizing bar plots (by proportion)
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(y = homeownership, fill = grade)) + geom_bar(position =
"fill") +labs( x ="Proportion", y ="Homeownership", fill ="Grade", title ="Grades of Lending Club loans", subtitle="and homeownership of lendee")
```

## Creating violin plots 
```{r, eval=TRUE, echo=TRUE}
ggplot(loans, aes(x = homeownership, y = loan_amount)) +
geom_violin()
```

## Creating ridge plots 
```{r, echo=TRUE, eval=TRUE}
library(ggridges)
ggplot(loans, aes(x = loan_amount, y = grade, fill = grade, color = grade)) +geom_density_ridges(alpha =0.5)
```

Thank you! 
