---
pdf_document: default
author: "Ariel Quek Xuan"
date: "`r Sys.Date()`"
output:
  html_document: default
  df_print: paged
  pdf_document: default
title: 'Week-4: Code-along_Ariel_Quek_Xuan'
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

# II. Code to edit and execute using the Code-along.Rmd file 
<br>

## A. Data Wrangling
<br>

### 1. Loading packages (Slide #16)

```{r, eval=TRUE,echo=TRUE}
# Load package tidyverse
install.packages("tidyverse",repos = "http://cran.us.r-project.org")
library(tidyverse)
```

### 2. Loading data-set (Slide #16)

```{r, eval=TRUE,echo=TRUE}
# Read data from the hotels.csv file and assign it to a variable named, "hotels"
hotels <- read.csv("hotels.csv")
head(hotels)
```

### 3. List names of the variables in the data-set (Slide #19)

```{r, eval=TRUE,echo=TRUE}
# Listing names of the variables in data-set (hotels)
names_hotels <- names(hotels)
head(names_hotels)
```


### 4. Glimpse of contents of the data-set (Slide #20)

```{r, eval=TRUE,echo=TRUE}
# to procure a glimpse of contents of the data-set
glimpse(hotels)
```
<br>

## B. Choosing rows or columns
<br>

### 5. Select a single column (Slide #24)

```{r, eval=TRUE,echo=TRUE}
# selecting a column
column <- select(hotels,lead_time)
head(column)
```

### 6. Select multiple columns (Slide #25)

```{r, eval=TRUE,echo=TRUE}
# selecting columns "lead_time", "agent", "market_segment"
columns <- select(hotels,lead_time,agent,market_segment)
head(columns)
```

### 7. Arrange entries of a column (Slide #28)

```{r, eval=TRUE,echo=TRUE}
# arrange by lead_time
arrangement <- arrange(hotels,lead_time)
head(arrangement)
```

### 8. Arrange entries of a column in the descending order (Slide #30)

```{r, eval=TRUE,echo=TRUE}
# arranging entries of "lead_time" in descending order
lead_time_arrangement <- arrange(hotels,desc(lead_time))
head(lead_time_arrangement)
```

### 9. Select columns and arrange the entries of a column (Slide #31)

```{r, eval=TRUE,echo=TRUE}
# selecting and arranging "lead_time" 
lead_time_arrangement_2 <- select(hotels,lead_time)
lead_time_arrangement_3 <- arrange(select(hotels,lead_time),desc(lead_time))
head(lead_time_arrangement_3)
```

### 10. Select columns and arrange the entries of a column using the pipe operator (Slide #37)

```{r, eval=TRUE,echo=TRUE}
df_1 <- hotels%>%
  select(lead_time)%>%
  arrange(desc(lead_time))
head(df_1)
```

### 11. Pick rows matching a condition (Slide #44)

```{r, eval=TRUE,echo=TRUE}
df_2 <- hotels %>% 
  filter(children >=1) %>% 
  select(hotel,children)
head(df_2)
```

### 12. Pick rows matching multiple conditions (Slide #46)

```{r, eval=TRUE,echo=TRUE}
df_3 <- hotels %>% 
  filter(children >=1,hotel=="City Hotel")%>%
  select(hotel,children)
head(df_3)
```

### 13. Non-conditional selection of rows: sequence of indices (Slide #49)

```{r, eval=TRUE,echo=TRUE}
hotels %>% 
  slice(1:5)
```

### 14. Non-conditional selection of rows: non-consecutive/specific indices (Slide #50)

```{r, eval=TRUE,echo=TRUE}
hotels %>% slice(1,3,5)
```

### 15. Pick unique rows using distinct() (Slide #52)

```{r, eval=TRUE,echo=TRUE}
hotels %>% 
  distinct(hotel)
```
<br>

## C. Creating new columns
<br>

### 16. Creating a single column with mutate() (Slide #56)

```{r, eval=TRUE,echo=TRUE}
single_column <- hotels%>%
  mutate(little_ones=children+babies)%>%
  select(hotel,little_ones,children,babies)
head(single_column)
```

### 17. Creating multiple columns with mutate() (Slide #58)

```{r, eval=TRUE,echo=TRUE}
multiple_columns <- hotels%>%
  mutate(little_ones=children+babies,average_little_ones=mean(little_ones))%>%
  select(hotel,little_ones,children,babies,average_little_ones)
head(multiple_columns)
```
<br>

## D. More operations with examples
<br>

### 18. count() to get frequencies (Slide #60)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  count(market_segment)
```

### 19. count() to get frequencies with sorting of count (Slide #61)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  count(market_segment,sort=TRUE)
```

### 20. count() multiple variables (Slide #62)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  count(hotel,market_segment)
```

### 21. summarise() for summary statistics (Slide #63)

```{r, eval=TRUE,echo=TRUE}
#mean average daily rate for all bookings 
hotels%>%
  summarise(mean_adr=mean(adr))
```

### 22. summarise() by using group_by to find mean (Slide #64)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  group_by(hotel)%>%
  summarise(mean_adr=mean(adr))
```

### 23. summarise() by using group_by to get count (Slide #65)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  group_by(hotel)%>%
  summarise(count=n())
```

### 24. summarise() for multiple summary statistics (Slide #67)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  summarise(min_adr=min(adr),mean_dr=mean(adr),median_adr=median(adr),max_adr=max(adr))
```

### 25. select(), slice() and arrange() (Slide #68)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  select(hotel,lead_time)%>%
  slice(1:5)%>%
  arrange(lead_time)
```

### 26. select(), arrange() and slice() (Slide #69)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  select(hotel,lead_time)%>%
  arrange(lead_time)%>%
  slice(1:5)
```

### 27. filter() to select rows based on conditions (Slide #73)

```{r, eval=TRUE,echo=TRUE}
filter1 <- hotels%>%
  filter(adults==0,children>=1)%>%
  select(adults,babies,children)
head(filter1)
```

### 28. filter() to select rows based on complicated conditions (Slide #74)

```{r, eval=TRUE,echo=TRUE}
filter2 <- hotels%>%
  filter(adults==1,children>=1|babies>=1)%>%
  select(adults,babies,children)
head(filter2)
```

### 29. count() and arrange() (Slide #76)

```{r, eval=TRUE,echo=TRUE}
hotels%>%
  count(market_segment)%>%
  arrange(desc(n))
```

### 30. mutate(), select() and arrange() (Slide #77)

```{r, eval=TRUE,echo=TRUE}
mutation1 <- hotels%>%
  mutate(little_ones=children+babies)%>%
  select(children,babies,little_ones)%>%
  arrange(desc(little_ones))
head(mutation1)
```


### 31. mutate(), filter() and select() (Slide #78)

```{r, eval=TRUE,echo=TRUE}
mutation2 <- hotels%>%
  mutate(little_ones=children+babies)%>%
  filter(little_ones>=1,hotel=="Resort Hotel")%>%
  select(hotel,little_ones)
head(mutation2)
```

```{r,eval=TRUE,echo=TRUE}
mutation3 <- hotels%>%
  mutate(little_ones=children+babies)%>%
  filter(little_ones>=1,hotel=="City Hotel")%>%
  select(hotel,little_ones)
head(mutation3)
```

Thank you! 