---
title: "Challenge-6"
author: "Ariel Quek Xuan"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Questions
<br>

#### Question-1: Countdown Blastoff (While Loop)

Create a program that simulates a rocket launch countdown using a while loop. Start from 10 and countdown to "Blastoff!" with a one-second delay between each countdown number. Print a message when the rocket launches.

**Hint:** You may want to use `cat` command to print the countdown and `Sys.sleep` for incorporating the delay

**Solutions:**

```{r, eval=TRUE,echo=TRUE}
# Creating a function for countdown 
rocket_countdown <- function() {
  countdown <- 10 
  while(countdown>0) {
    cat(countdown, "...","")
    countdown <- countdown - 1
    Sys.sleep(1)
  }
  cat("Blast off!")
}

rocket_countdown()
```
<br>

#### Question-2: Word Reverser (for Loop)

Develop a program that takes a user-entered word and uses a while loop to print the word's characters in reverse order. For example, if the user enters "hello," the program should print "olleh."

**Hint:** You may want to use `substr` command to access each character of the input word, and `paste` command to join the reversed letters one at a time

**Solutions:** 

```{r, eval=TRUE,echo=TRUE}
Word_reverser <- function() {
  user_input <- readline("Enter your word here:")
  reversed_word <- "" 
  for (i in nchar(user_input):1) {
  reversed_word <- paste(reversed_word, substr(user_input,i,i),sep="")
}
cat("This is your reversed word:",reversed_word)
}
Word_reverser()

```

<br>

Thank you! Have a wonderful recess week! 
